---
title: "Boston Housing Prices"
output: html_notebook
---

Camron Khan<br>
April 2017

## Dependencies
```{r}
#install.packages("caret")
#install.packages("ggplot2")
#install.packages("AppliedPredictiveModeling")

library(caret)
library(ggplot2)
library(reshape2)
library(AppliedPredictiveModeling)
```

## Data Source
```r
rawData <- read.csv('./Boston.csv', stringsAsFactors = FALSE)
```
```{r}
rawData <- read.csv('./Boston.csv', stringsAsFactors = FALSE)
```

## Definitions

|Column   |Type   |Definition                                                               |
|---------|-------|-------------------------------------------------------------------------|
|X        |int    |index                                                                    |
|crim     |real   |per capita crime rate by town                                            |
|zn       |real   |proportion of residential land zoned for lots over 25,000 sq.ft          |
|indus    |real   |proportion of non-retail business acres per town                         |
|chas     |int    |Charles River dummy variable (= 1 if tract bounds river; 0 otherwise)    |
|nox      |real   |nitrogen oxides concentration (parts per 10 million)                     |
|rm       |real   |average number of rooms per dwelling                                     |
|age      |real   |proportion of owner-occupied units built prior to 1940                   |
|dis      |real   |weighted mean of distances to five Boston employment centers             |
|rad      |int    |index of accessibility to radial highways                                |
|tax      |int    |full-value property-tax rate per \$10,000                                |
|ptratio  |real   |pupil-teacher ratio by town                                              |
|black    |real   |1000(Bk - 0.63)^2 where Bk is the proportion of blacks by town           |
|lstat    |real   |lower status of the population (percent)                                 |
|medv     |real   |median value of owner-occupied homes in \$1000s (**Dependent Variable**) |

## Data Summary
like this
```{r}
str(rawData)
```


```{r}
#ggplot(melt(DATA), aes(x=value)) + geom_histogram() + facet_wrap(~variable)

qplot(rawData$rm, geom = 'histogram', binwidth = .1) + xlab('')
qplot(rawData$crim, geom = 'histogram', binwidth = 1) + xlab('')

```





```{r}
preprocessor = preProcess(rawData[2:15], c("BoxCox", "center", "scale"))
preprocessedData = data.frame(predict(preprocessor, rawData[2:15]))
```




```{r}
cor(rawData[,2:15])
```


```{r}
cor(preprocessedData)
```



```{r, fig.width=15, fig.height=10}
ggplot(melt(preprocessedData), aes(x=value)) + geom_histogram(binwidth = .5) + facet_wrap(~variable)
```



